---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  cache = TRUE
)
stepcount::unset_reticulate_python()
stepcount::use_stepcount_condaenv()
library(stepcount)
library(reticulate)
options(digits.secs = 3)
```

# stepcount

<!-- badges: start -->
[![R-CMD-check](https://github.com/jhuwit/stepcount/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhuwit/stepcount/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of `stepcount` is to wrap up the https://github.com/OxWearables/stepcount algorithm.

# Installation

## Install `stepcount` Python Module

See https://github.com/OxWearables/stepcount?tab=readme-ov-file#install for how to install the `stepcount` python module.

### Issues
If you are using the Random Forest model from `stepcount`, you may need `hmmlearn<0.3.0` due to some issues with its new implementation of its models as described https://github.com/OxWearables/stepcount/issues/62 (Feb 2024).

## Install `stepcount` R Package

You can install the development version of `stepcount` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jhuwit/stepcount")
```


# Usage

## Loading the `stepcount` conda environment 
In order to use the `stepcount` conda environment that OxWearables recommends, you must run this command **before `reticulate`** is loaded:
```{r eval = FALSE}
stepcount::use_stepcount_condaenv()
```

## The `RETICULATE_PYTHON` environment variable
If you have the `RETICULATE_PYTHON` environment variable in your `.Renviron` or your `PATH`, then `reticulate` will still use that version of Python and the code will likely not work.  The `unset_reticulate_python()` function will unset that environment variable.  So the usage would start with something like:

```{r eval = FALSE}
stepcount::unset_reticulate_python()
stepcount::use_stepcount_condaenv()
```

and if you need `reticulate`, you would load it after

```{r eval = FALSE}
stepcount::unset_reticulate_python()
stepcount::use_stepcount_condaenv()
library(reticulate)
```

The `stepcount_check` function can determine if the `stepcount` module can be loaded
```{r stepcount_check}
stepcount::stepcount_check()
```

## Running `stepcount` (file)
The main function is `stepcount::stepcount`, which takes can take in a file directly:

```{r run_stepcount_file}
library(stepcount)
library(dplyr)
library(ggplot2)
library(tidyr)
file = system.file("extdata/P30_wrist100.csv.gz", package = "stepcount")
if (stepcount_check()) {
  out = stepcount(file = file)
}
```

Let's see inside the output, which is a list of values, namely a `data.frame` of `steps` with the time (in 10s increments) and the number of steps in those 10 seconds, a `data.frame` named `walking` which has indicators for if there is walking within that 10 second period:

```{r stepcount_output}
names(out)
str(out)
head(out$steps)
tail(out$steps)
tail(out$walking)
```

The `step_times` `data.frame` indicates which times are when steps occurred (at the original sample rate).  Make sure you have the `digits.secs` option set to see the sub-seconds for the times (esp for writing out files in `readr::write_csv`):

```{r stepcount_output_steptimes}
head(out$step_times)
options(digits.secs = 3)
head(out$step_times)
```

We can plot a portion of the tri-axial data and show where the steps were indicated:
```{r stepcount_plot}
df = readr::read_csv(file)
if (stepcount_check()) {
  st = out$step_times
    dat = df[10000:12000,] %>%
      dplyr::select(-annotation) %>%
      tidyr::gather(axis, value, -time)
    st = st %>%
      dplyr::mutate(time = lubridate::as_datetime(time)) %>%
      dplyr::as_tibble()
    st = st %>%
      dplyr::filter(time >= min(dat$time) & time <= max(dat$time))
    dat %>%
      ggplot2::ggplot(ggplot2::aes(x = time, y = value, colour = axis)) +
      ggplot2::geom_line() +
      ggplot2::geom_vline(data = st, ggplot2::aes(xintercept = time))
}
```

The main caveat is that `stepcount` is very precise in the format of the data, primarily it must have the columns `time`, `x`, `y`, and `z` in the data.  

## Running `stepcount` (data frame)
Alternatively, you can pass out a `data.frame`, rename the columns to what you need them to be and then run `stepcount` on that:

```{r run_stepcount_df}
head(df)
out_df = stepcount(file = df)
```

Which gives same output for this data:

```{r stepcount_file_df_check, dependson=c("run_stepcount_file", "run_stepcount_df")}
all.equal(out[c("steps", "walking", "step_times")],
          out_df[c("steps", "walking", "step_times")])
```


## Running `stepcount` on multiple files
When you pass in multiple files, `stepcount` will run all of them, but it will only load the model once, which can have savings, but the results are still in memory:
```{r multifile}
if (stepcount_check()) {
  out2 = stepcount(file = c(file, file))
  length(out2)
  names(out2)
  # all.equal(out[c("steps", "walking", "step_times")], 
  #           out2[[1]][c("steps", "walking", "step_times")])
}
```


## Stepcount Random Forest
The `model_type` parameter indicates the type of model being run, and the `rf` will provide the predictions from a random forest

```{r run_stepcount_rf, eval = FALSE}
if (stepcount_check()) {
  out_rf = stepcount(file = file, model_type = "rf")
}
```

